---
title: 'Home'
output:
  distill::distill_article:
    toc:false:
      theme: theme.css
favicon: "images/favicon.ico"
github-repo: childpovertyactionlab/safmr
---

```{r Library Loading and Path Set-Up, include = FALSE}
#### Libraries to load #####
library(tidyverse)
library(sf)
library(leaflet)
library(mapboxapi)
library(rio)
library(CPALtools)

libDB <- "C:/Users/micha/CPAL Dropbox/" # Michael Laptop
#libDB <- "E:/CPAL Dropbox/" # Michael Desktop

```

```{r, Data and File Import, include = FALSE}
# simple boundary layer of the city of dallas
dallas <- st_read("data/dallas_simpleboundary.geojson") %>%
  st_transform(crs = 4326)

# layer of zip codes with safmr and average asking rent in city of dallas
safmr <- st_read("data/safmr_zcta.geojson") %>%
  st_transform(crs = 4326) %>%
  filter(!is.na(medRent2BR)) %>%
  st_intersection(., dallas)
```

## Small Area Fair Market Rent Difference from Average Rent

This map is generated using 

* Small Area Fair Market Rent estimates from the Department of Housing and Urban Development for 2022
* Two-Bedroom average rent for multi-family apartment complexes from CoStar for Q1 2022

CoStar data is aggregated to the zip code level and averaged for all complexes within the boundaries of the City of Dallas.
The 'Rent Difference' metric calculates the difference between SAFMR and average rent for a zip code.

CoStar tracks properties with five or more units, so data on smaller properties with 1-4 units is not reflected here.

```{r Leaflet Map, echo = FALSE, message=FALSE, warning=FALSE, layout = "l-body-outset", fig.height=10}
#names(safmr)
#min(safmr$RentDiff, na.rm = TRUE)
#max(safmr$RentDiff, na.rm = TRUE)

rent.labels <- c("Less than SAFMR",
                 "Up to $250 more than SAFRM",
                 "More than $250 over SAFRM")

rent.bins <-c(-500, 0, 250, 1500)
rent.pal <- colorBin(palette = pal_cpaltri, 
                     domain = safmr$RentDiff, 
                     bins = rent.bins, 
                     na.color = "#595959")

bbox <- st_bbox(dallas) %>%
  as.vector()

popupRent <- paste0("<b>", "Zip Code: ", "</b>", safmr$GEOID, "<br>",
                "<b>", "Multi-Family Properties: ", "</b>", safmr$TotMF, "<br>",
                "<b>", "Fair Market Rent 2-Bedroom: ", "</b>", "$", safmr$FairMrkt2BR, "<br>",
                "<b>", "Average Rent 2-Bedroom: ", "</b>", "$", safmr$medRent2BR, "<br>",
                "<b>", "Rent Difference: ", "</b>", "$", safmr$RentDiff, "<br>"
                )

leaflet() %>%
  fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>%
  addTiles(urlTemplate = cpal_mapbox, attribution = cpal_leaflet) %>%
  addPolygons(data = safmr,
              fillColor = ~rent.pal(RentDiff),
              weight = 1,
              opacity = 0.8,
              color = "#FFFFFF",
              fillOpacity = 0.7,
              popup = ~popupRent,
              highlight = highlightOptions(
                weight = 4,
                color = "#FFFFFF",
                fillOpacity = 0.4,
                bringToFront = TRUE)) %>%
    addLegend(data = safmr,
            pal = rent.pal, 
            values = ~rent.bins, 
            opacity = 1, 
            title = "SAFRM to Average Rent Difference",
            position = "bottomright",
            labFormat = function(type, cuts, p){  # Here's the trick
                                                  paste0(rent.labels)
                             })
```