---
title: 'SAFMR to MAR'
---

```{r Library Loading and Path Set-Up}
#| include: false
#### Libraries to load #####
library(tidyverse)
library(sf)
library(leaflet)
library(mapboxapi)
library(rio)
library(cpaltemplates)

ntxcounty <- st_read("data/ntx_county.geojson") %>%
  st_transform(crs = 4326)

ntxDissolve <- ntxcounty %>%
  st_union(.)

# layer of zip codes with safmr and average asking rent in city of dallas
safmr <- st_read("data/safmr_zcta.geojson") %>%
  st_transform(crs = 4326)
```

This map is generated using 

* Small Area Fair Market Rent from the Department of Housing and Urban Development for FY 2024 for two-bedroom unit
* Two-Bedroom average rent for multi-family apartment complexes from CoStar for Q3 2023

CoStar data is aggregated to the zip code level and averaged for all complexes within the boundaries North Texas.
The 'Rent Difference' metric calculates the difference between average rent for a zip code and SAFMR.

CoStar tracks properties with five or more units, data on smaller properties with 1-4 units is not reflected here.

```{r Leaflet Map}
#| column: screen-inset-shaded
#| fig-height: 12
rent.labels <- c("Less than SAFMR",
                 "Up to $250 more than SAFMR",
                 "More than $250 over SAFMR")

rent.bins <-c(-1500, 0, 250, 1000)
rent.pal <- colorBin(palette = c(palette_cpal_main[1], palette_cpal_main[4], palette_cpal_main[2]), 
                     domain = safmr$RentDiff, 
                     bins = rent.bins, 
                     na.color = "#595959")

bbox <- st_bbox(ntxDissolve) %>%
  as.vector()

popupRent <- paste0("<b>", "Zip Code: ", "</b>", safmr$ZCTA, "<br>",
                "<b>", "Multi-Family Properties: ", "</b>", safmr$TotMF, "<br>",
                "<b>", "Total Units: ", "</b>", safmr$TotUnits, "<br>",
                "<b>", "Fair Market Rent 2-Bedroom: ", "</b>", "$", safmr$fairMrkt2BR, "<br>",
                "<b>", "Average Rent 2-Bedroom: ", "</b>", "$", safmr$medRent2BR, "<br>",
                "<b>", "Rent Difference: ", "</b>", "$", safmr$RentDiff, "<br>"
                )

leaflet() %>%
  fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>%
  addTiles(urlTemplate = cpal_mapbox, attribution = cpal_leaflet) %>%
  addPolygons(data = safmr,
              fillColor = ~rent.pal(RentDiff),
              weight = 1,
              opacity = 0.8,
              color = "#FFFFFF",
              fillOpacity = 0.7,
              popup = ~popupRent,
              highlight = highlightOptions(
                weight = 4,
                color = "#FFFFFF",
                fillOpacity = 0.2,
                bringToFront = TRUE)) %>%
    addLegend(data = safmr,
            pal = rent.pal, 
            values = ~rent.bins, 
            opacity = 1, 
            title = "Market Asking Rent is:",
            position = "bottomright",
            labFormat = function(type, cuts, p){  # Here's the trick
                                                  paste0(rent.labels)
                             })
```